---
title: "Covid19 vaccine study Fall2021"
author: "Sierra Green"
date: "`12/03/2021`"
output: 
  html_document:
    toc: true
    toc_depth: 5
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
TheData <-read.csv(file="FallFinalData.csv",header=TRUE) %>% select(-"X.1", -"X")
Covid91 <- TheData %>% mutate(across(where(is.character),as.factor)) 
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males
```{r}
Males<- filter(Covid91,sex == "Man")
```

```{r}
barchartGC(~treatment + infected,data=Males)
barchartGC(~treatment + infected,data=Males, type="percent")

```


The bar chart shows us that those who took the placebo got sick more than the males who took the Covid91. We can see that the pink bar for the placebo is almost double in size compared to the drug. This indicates is effective for the males group.



## Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```



33.11% of sick males had the treatment, and 66.89% of males got sick when they took the placebo. This is a 2:1 difference and lets us know that the treatment is actually working.



## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Malemod <- fisher.test(table1)
```



According to the Fisher's test, the p-value is extremely small. This proves we would have to reject the null hypothesis. There is difference between the placebo and those who took the drug.



```{r}
MaleEstimate <- Malemod$estimate
Malelow <- Malemod$conf.int[1]
Malehigh <- Malemod$conf.int[2]
Malelowprob <- Malelow / 1+ Malelow
Maleloweffect <- Malelow * 100
Malehighprob <- Malehigh / 1+ Malehigh
Malehigheffect <- Malehigh * 100
Maleprob <- MaleEstimate / 1+ MaleEstimate
MaleEffect <- Maleprob * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccine are `r MaleEffect`. The confidence interval for males is `r Maleloweffect` to `r Malehigheffect`.

# Females

```{r}
Females <- filter(Covid91, sex == "Woman")
```


```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```


According to the chart, it seems as though the vaccine is not going well for the females. The blue bar indicates that the females are getting sick than healthy; more than 50%.



## Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```


This states that 58% of females who took the drug got very sick and only 42% gotten sick from taking the placebo. The vaccine is cleary not taking effect for the females.






## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Femalemod <- fisher.test(table1)
```



The p-vale is again very small (6e-04). Therefore, we would consider rejecting the null hypothesis. This data took a turn in a very bad way. Majority of the females were getting sick.




```{r}
FemaleEstimate <- Femalemod$estimate
Femalelow <- Femalemod$conf.int[1]
Femalehigh <- Femalemod$conf.int[2]
Femalelowprob <- Femalelow / 1+ Malelow
Femaleloweffect <- Femalelow * 100
Femalehighprob <- Femalehigh / 1+ Femalehigh
Femalehigheffect <- Femalehigh * 100
Femaleprob <- FemaleEstimate / 1+ FemaleEstimate
FemaleEffect <- Femaleprob * 100
options(digits = 1)
```


The effectiveness of the Covid91 vaccine are `r FemaleEffect`. The confidence interval for males is `r Femaleloweffect` to `r Femalehigheffect`.


# Drug Users

# LGBTQ

# Conclusions and recommendations

