---
title: "Covid19 vaccine study Fall2021"
author: "Sierra Green"
date: "`12/03/2021`"
output: 
  html_document:
    toc: true
    toc_depth: 5
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
TheData <-read.csv(file="FallFinalData.csv",header=TRUE) %>% select(-"X.1", -"X")
Covid91 <- TheData %>% mutate(across(where(is.character),as.factor)) 
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males
```{r}
Males<- filter(Covid91,sex == "Man")
```

```{r}
barchartGC(~treatment + infected,data=Males)
barchartGC(~treatment + infected,data=Males, type="percent")

```


The bar chart shows us that those who took the placebo got sick more than the males who took the Covid91. We can see that the pink bar for the placebo is almost double in size compared to the drug. This indicates is effective for the males group.



## Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```



33.11% of sick males had the treatment, and 66.89% of males got sick when they took the placebo. This is a 2:1 difference and lets us know that the treatment is actually working.



## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Malemod <- fisher.test(table1)
```



According to the Fisher's test, the p-value is extremely small. This proves we would have to reject the null hypothesis. There is difference between the placebo and those who took the drug.



## Calculating Effectiveness

```{r}
MaleEstimate <- Malemod$estimate
Malelow <- Malemod$conf.int[1]
Malehigh <- Malemod$conf.int[2]
Malelowprob <- Malelow / 1+ Malelow
Maleloweffect <- Malelow * 100
Malehighprob <- Malehigh / 1+ Malehigh
Malehigheffect <- Malehigh * 100
Maleprob <- MaleEstimate / 1+ MaleEstimate
MaleEffect <- Maleprob * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccine are `r MaleEffect`. The confidence interval for males is `r Maleloweffect` to `r Malehigheffect`.

# Females

```{r}
Females <- filter(Covid91, sex == "Woman")
```


```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```


According to the chart, it seems as though the vaccine is not going well for the females. The blue bar indicates that the females are getting sick than healthy; more than 50%.



## Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```


This states that 58% of females who took the drug got very sick and only 42% gotten sick from taking the placebo. The vaccine is clearly not taking effect for the females.






## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Femalemod <- fisher.test(table1)
```



The p-vale is again very small (6e-04). Therefore, we would consider rejecting the null hypothesis. This data took a turn in a very bad way. Majority of the females were getting sick.


## Calculating Effectiveness

```{r}
FemaleEstimate <- Femalemod$estimate
Femalelow <- Femalemod$conf.int[1]
Femalehigh <- Femalemod$conf.int[2]
Femalelowprob <- Femalelow / 1+ Femalelow
Femaleloweffect <- Femalelow * 100
Femalehighprob <- Femalehigh / 1+ Femalehigh
Femalehigheffect <- Femalehigh * 100
Femaleprob <- FemaleEstimate / 1+ FemaleEstimate
FemaleEffect <- Femaleprob * 100
options(digits = 1)
```


The effectiveness of the Covid91 vaccine are `r FemaleEffect`. The confidence interval for males is `r Femaleloweffect` to `r Femalehigheffect`.


# Drug Users


```{r}
DrugUsers <- filter(Covid91, DrugUser == "yes")
```


```{r}
barchartGC(~infected + treatment,data=DrugUsers)
barchartGC(~infected + treatment,data=DrugUsers, type="percent")

```


The bar chart shows that majority of the drug users who consumed the placebo became very sick rather than healthy however the people who took the drug were more healthy than they were sick.

## Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=DrugUsers)
rowPerc(table1)
colPerc(table1)
```


It seems that 48% of drug users who took the drug ended up not being sick leaving 33% sick whereas 52% remain healthy leaving 67% of drug users who consumed the placebo became sick. This shows that the placebo took more effect in a bad way for drugusers. 



## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Drugmod <- fisher.test(table1)
```


According to the Fisher's test, the p-value is small leaving it at 0.003. This indicates that we would have to reject the null hypothesis. However, drug users did get sick more than than being healthy when consuming the placebo.


## Calculating Effectiveness

```{r}
DrugEstimate <- Drugmod$estimate
Druglow <- Drugmod$conf.int[1]
Drughigh <- Drugmod$conf.int[2]
Druglowprob <- Druglow / 1+ Druglow
Drugloweffect <- Druglow * 100
Drughighprob <- Drughigh / 1+ Drughigh
Drughigheffect <-Drughigh * 100
Drugprob <- DrugEstimate / 1+ DrugEstimate
DrugEffect <- Drugprob * 100
options(digits = 1)
```




The effectiveness of the Covid91 vaccine are `r DrugEffect`. The confidence interval for males is `r Drugloweffect` to `r Drughigheffect`.


# LGBTQ

```{r}
LGBTQ <- filter(Covid91, LGBTQ == "gay")
```

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")
```




According to the bar chart, it seems as though the placebo took affect in a good way rather than the regular drug. More people ended being healthy than sick when taking the placebo, however, when the regular drug was consumed, it shows that more people got sick from it.



## Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```


The table represents that 71% of people got sick leaving 51% of them healthy when the drug was consumed. When the placebo was consumed, it left 29% of them sick and 51% of them healthy. The placebo seemed to work.


## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
LGBTQmod <- fisher.test(table1)
```



The p-value is at a 0 which means that we would reject the null hypothesis. The placebo did work for majority of the LGBTQ people.


## Calculating Effectivenss

```{r}
LGBTQEstimate <- LGBTQmod$estimate
LGBTQlow <- LGBTQmod$conf.int[1]
LGBTQhigh <- LGBTQmod$conf.int[2]
LGBTQlowprob <- LGBTQlow / 1+ LGBTQlow
LGBTQloweffect <- LGBTQlow * 100
LGBTQhighprob <- LGBTQhigh / 1+ LGBTQhigh
LGBTQhigheffect <-LGBTQhigh * 100
LGBTQprob <- LGBTQEstimate / 1+ LGBTQEstimate
LGBTQEffect <- LGBTQprob * 100
options(digits = 1)
```


The effectiveness of the Covid91 vaccine are `r LGBTQEffect`. The confidence interval for males is `r LGBTQloweffect` to `r LGBTQhigheffect`.



# Conclusions and recommendations